Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2025-02-21 paper.tex
Appendix A

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
Except for a subset of initial conditions, the solutions to the fully nonlinear
crest-action conservation equations are mostly steady in the reference frame of
the long waves.
\end{abstract}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
\caption{
  A diagram of short waves riding on longer waves that propagate with their
  phase-speed $C_{pL}$ from left to right.
  The short waves propagate on the long-wave surface, their action density
  moving with the group velocity $C_g$.
  Long waves induce orbital velocities $U$ that cause near-surface convergence
  and divergence on their front and rear faces, respectively.
  They also induce downward and upward centripetal accelerations
  ($\partial W/\partial t$) in the crests and troughs, respectively, that
  modulate the effective gravity at the surface.
  As a consequence of surface convergence and divergence and the effective-gravity
  modulation, the short waves become shorter and higher preferentially
  on the long-wave crests.
  The diagram is not to scale.
}
\label{fig:hydrodynamic_modulation_diagram}
\end{figure}

@@

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action density of short waves defined as the ratio of their
energy density to their intrinsic frequency, $E/\sigma$,
and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
The wave action density expressed as $E/\sigma$ is commonly taken as
conservative in wave modulation theory
\citep{phillips1981dispersion,longuet1987propagation}, however as
\citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
Although the absence of wave growth and dissipation is not generally realistic in
the open ocean, it is a useful approximation here to isolate the effects
of hydrodynamic modulation solely due to the motion of the long waves.

@@

\subsection{Limits to the wave action conservation}
\label{appendix:wave_action_conservation}

@@

@@

The long-wave induced velocities (\ref{eq:phi})-(\ref{eq:W_L}) are accurate to
$\mathcal{O}(\varepsilon_L^3)$ when evaluated at the mean water level.
However, in this paper we depart from the mean water level and evaluate the
velocities at the surface of the long wave.
This introduces an additional error at the order of $\mathcal{O}(\varepsilon_L^2)$.
To quantify this error exactly, we compare the velocities evaluated at the
surface for a linear (small-amplitude) wave and a third-order Stokes wave,
with the surface velocity of a fully nonlinear wave.
The nonlinear wave surface velocity is calculated numerically using the method
by \citet{clamond2018accurate}.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_phase.pdf}
\caption{
}
\label{fig:velocity_error_by_phase}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_ak.pdf}
\caption{
}
\label{fig:velocity_error_by_ak}
\end{figure}

>>>>

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
Except for a subset of initial conditions, the solutions to the fully nonlinear
crest-action conservation equations are mostly steady in the reference frame of
the long waves.
\end{abstract}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
\caption{
  A diagram of short waves riding on longer waves that propagate with their
  phase-speed $C_{pL}$ from left to right.
  The short waves propagate on the long-wave surface, their action
  moving with the group velocity $C_g$.
  Long waves induce orbital velocities $U$ that cause near-surface convergence
  and divergence on their front and rear faces, respectively.
  They also induce downward and upward centripetal accelerations
  ($\partial W/\partial t$) in the crests and troughs, respectively, that
  modulate the effective gravity at the surface.
  As a consequence of surface convergence and divergence and the effective-gravity
  modulation, the short waves become shorter and higher preferentially
  on the long-wave crests.
  The diagram is not to scale.
}
\label{fig:hydrodynamic_modulation_diagram}
\end{figure}

@@

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action of short waves defined as the ratio of their
energy density to their intrinsic frequency, $E/\sigma$,
and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
The wave action expressed as $E/\sigma$ is commonly taken as
conservative in wave modulation theory
\citep{phillips1981dispersion,longuet1987propagation}, however as
\citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
We describe this requirement on gentle wave slope in more detail further below.
Although the absence of wave growth and dissipation is not generally realistic in
the open ocean, it is a useful approximation here to isolate the effects
of hydrodynamic modulation solely due to the motion of the long waves.

@@

\subsection{Limits to the wave action conservation}
\label{sec:wave_action_conservation}

@@

As the numerical solutions are for nonlinear wave action and crest conservation
equations, it is important to re-evaluate the homogeneity and stationarity
requirements for the wave action conservation to remain valid.
Fig. \ref{fig:homogeneity_stationarity_numerical} shows the homogeneities
and stationarities of the short-wave wavenumber, action, and effective gravity
as function of the long-wave steepness $\varepsilon_L$ and the wavenumber ratio
$k/k_L$, based on the numerical simulations.
Although homogeneity is slightly lower ($H_{k,N,g} \approx 0.92$ for $k/k_L = 10$
and $\varepsilon_L = 0.4$) compared to the analytical solutions,
the stationarity is significantly reduced.
Following the same criteria for strong ($S_{k,N,g} > 0.99$) and weak
($S_{k,N,g} > 0.9$) stationarity as in subsection \ref{sec:wave_action_conservation},
the numerical solutions are only weakly stationary for $\varepsilon_L < 0.21$
at $k/k_L = 10$, and for $\varepsilon_L < 0.36$ at $k/k_L = 100$.
This provides a quantitative guidance on when the short-wave action can be
considered to be conserved.

@@

The long-wave induced velocities (\ref{eq:phi})-(\ref{eq:W_L}) are accurate to
$\mathcal{O}(\varepsilon_L^3)$ when evaluated at the mean water level.
However, in this paper we depart from the mean water level and evaluate the
velocities at the surface of the long wave.
This introduces a additional errors at $\mathcal{O}(\varepsilon_L^2)$.
To quantify this error exactly, we compare the velocities evaluated at the
surface for a linear (small-amplitude) wave and a third-order Stokes wave,
with the surface velocity of a fully nonlinear wave
(Fig. \ref{fig:velocity_error_by_phase}).
The elevation and the surface velocity for the fully nonlinear wave is
calculated numerically using the method by \citet{clamond2018accurate},
which relies on Petviashvili's iterations and in this case uses 2048 Fourier
modes to represent the wave.
Both the 1$^{st}$ and the 3$^{rd}$ order waves underestimate the surface velocity
at the crest and in the trough, and overestimate it on the front and rear faces
of the crest.

Maximum values of surface velocity errors for both the linear and the
3$^{rd}$ order Stokes wave asymptote to $\varepsilon_L^3$ for small
$\varepsilon_L$ (Fig. \ref{fig:velocity_error_by_ak}).
However, they approach and exceed $\varepsilon_L^2$ at $\varepsilon_L \approx 0.3$
and beyond.
Although the maximum error is $\approx 10-20\%$ smaller for the 3$^rd$ order
Stokes wave relative to the linear wave, its average error is slightly larger.
The main takeaway from this figure is that evaluating the surface velocity
of a linear wave carries an error no greater than $\varepsilon_L^2$ for
$\varepsilon_L < 0.32$.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_velocity_error_by_phase.pdf}
\caption{
  Surface elevation (top) and horizontal velocity (middle) for the linear wave
  (blue), the 3$^{rd}$ order Stokes wave (orange), and the fully nonlinear wave
  (black) with steepness of $\varepsilon_L = 0.4$.
  The bottom panel shows the surface velocity error for the linear wave (blue)
  and the 3$^{rd}$ order Stokes wave (orange), relative to the fully nonlinear
  wave solution.
}
\label{fig:velocity_error_by_phase}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_velocity_error_by_ak.pdf}
\caption{
  Surface velocity error as function of wave steepness $\varepsilon_L$,
  for the linear wave (blue) and the 3$^{rd}$ order Stokes wave (orange).
  Maximum and mean errors are indicated with solid and dashed lines,
  respectively.
  For reference, $\varepsilon_L^2$ and $\varepsilon_L^3$ curves are shown in
  black dashed and dotted lines, respectively.
}
\label{fig:velocity_error_by_ak}
\end{figure}
